---
title: "1129 HW6-R__Yan Yilin"
output: html_notebook
---

Question 1: Use R to create the following two matrices and do the indicated matrix multiplication.
What is the resulting matrix? (4 points)
```{r}
# Create the two required matrices
Q1MatrixA <- matrix(c(7,2,9,4,12,13), nrow = 2, ncol = 3)
Q1MatrixB <- matrix(c(1,2,3,7,8,9,12,13,14,19,20,21), nrow = 3, ncol = 4)

# Print the matrices
Q1MatrixA
Q1MatrixB
```
```{r}
# Do the indicated matrix multiplication
Q1Result = Q1MatrixA %*% Q1MatrixB
Q1Result
```
The resulting matrix is shown above.

Question 2: (total 13 points)
a. Create a data frame like the following table (2 points).
```{r}
# Create a data frame
Data_Frame <- data.frame (
  id = c(1,2,3,4,5),
  name = c("Peter", "Amy", "Ryan","Gary","Michelle"),
  salary = c(623.3, 515.2, 611.00,729.00, 843.25)
)

# Print the data frame
Data_Frame
```
b. Then add a column of “department”, and assign department names for the 5 persons, for example, IT, finance, computer science, etc., name your own. (2 points).
```{r}
# Add the "department" column
New_col_DF <- cbind(Data_Frame, department = c("IT","Finance","Accounting","Computer Science","Public Relation"))

#Print the new column
New_col_DF
```
c. Extract 1st, 3rd, and 5th rows, with 2nd and 3rd columns, using c function. (3 points).
```{r}
Extract_result = New_col_DF[c(1,3,5),c(2,3)]
Extract_result
```
d. plot a bar chart to show the salary of row 1,4,5 and label the bar with people’s name correspondingly. (3 points).
```{r}
# x-axis values: people's name
x <- New_col_DF[c(1,4,5),c(2)]

# y-axis values: salary
y <- New_col_DF[c(1,4,5),c(3)]

barplot(y, names.arg = x)
```
e. plot a pie chart, based on this data framework, only showing the highest salary, lowest salary, and median salary, give your plot a title, and label each pie. (3 points).
```{r}
H_S <- max(New_col_DF$salary)
H_L <- min(New_col_DF$salary)
H_M <- median(New_col_DF$salary)
H_S
H_L
H_M

# Create a vector of labels
mylabel <- c("Highest Salary", "Lowest Salary", "Median Salary")

# Create a vector of colors
colors <- c("palegreen", "paleturquoise", "papayawhip")

x <- c(H_S,H_L,H_M)

# Display the pie chart with colors
pie(x, label = mylabel, main = "Salary Pie Chart", col = colors)

# Display the explanation box
legend("bottomleft", mylabel, fill = colors)
```
Question 3: Pick two functions in your final project, transfer that into R programming. One function should involve “if else” (4 points), the other function should involve “for loop or while loop” (4 points). 
•	Test your function, showing your function works. You can assign some values, in order to test the function. 
•	This is individual homework, so you and your team member may pick the same functions but could have different translations. 
•	Note, in your homework submission, paste your python code with your R code for comparison. 
1) Transfer one "if else" function:
```{r}
# Note that by the previous assumptions and settings, ai_win and player_win won't be >2. ai_win or player_win can only be 0,1,2
# So the value assigned here (for the testing purpose) should be numbers from 0,1,2
ai_win = 2
player_win = 1

if (ai_win == 2) {
  print("You LOSE the game since ai has already won 2 rounds.")
} else{
    print("You've won 2 rounds, You WIN the game, Congratulation!!!")
  }
```
2) Transfer one "while loop" function:
```{r}
# by previous setting: replay = input("Play again? (yes/no)")
replay = "noo"

while (replay != "yes" && replay != "no"){
  print("The input is not valid, it should be one of 'yes' or 'no', please retype.")
  replay <- readline(prompt ="Play again? (yes/no)" )
}
```

